The Effect of a Second AIDS Defining Illness on Survival in the MACS

### 2015-05-10 paper.Rmd
Fix intervals again

# read in horiz data with lots of covariates
covariates <- read_csv("data/macswihs_horiz2013_copy-1.csv") %>%
           # date of death
    mutate(death.date=as.Date("1960-01-01") + FRSTDTHD,
           # date of aids diagnosis
           aids.date=as.Date("1960-01-01") + FRSTAIDD,
           # aids interval
           aids.interval=ifelse(year(aids.date) < 1990, "[84, 90)", 
                         ifelse(year(aids.date) < 1993, "[90, 93)",
                         ifelse(aids.date < as.Date("1995-07-01"), 
                                "[93, 07-95)",
                                "[07-95, 96)"))),
           # date of secondary aids diagnosis
           sec.aids.date=as.Date("1960-01-01") + SECAIDD,
           # sec aids interval
           seca.interval=ifelse(year(sec.aids.date) < 1990, "[84, 90)", 
                         ifelse(year(sec.aids.date) < 1993, "[90, 93)",
                         ifelse(sec.aids.date < as.Date("1995-07-01"), 
                                "[93, 07-95)",
                                "[07-95, 96)"))),
           # what was the first aids diagnosis
           frst.aids.diag=plyr::revalue(as.character(AIDSDIAG), map, 
                                        warn_missing=FALSE),
           # what was the first aids diagnosis
           sec.aids.diag=plyr::revalue(as.character(SECAIDSDIAG), map, 
                                       warn_missing=FALSE),
           # did the person have a sec aids diag
           had.sec=year(sec.aids.date) < 1996,
           dob=as.Date("1960-01-01") + DOB,
           age.at.aids=difftime(aids.date, dob, unit="weeks") / 52.25,
           age.at.aids=as.numeric(age.at.aids)) %>%
    select(STUDYID, 
           aids.date,
           aids.interval,
           sec.aids.date,
           seca.interval,
           had.sec,
           frst.aids.diag, 
           sec.aids.diag,
           death.date,
           race=RACE,
           age.at.aids) %>%
    # only keep pts who developed aids before 1996
    filter(year(aids.date) < 1996) %>%
    mutate(aids.year=year(aids.date),
           aids.month=month(aids.date),
           race=ifelse(race == 1, "white", "non-white")) %>%
    rowwise() %>%
    mutate(sec.aids.date=min(sec.aids.date, 
                             as.Date("1995-12-31"))) %>%
    ungroup() %>%
    mutate(sec.aids.date=ifelse(had.sec, sec.aids.date, 
                                as.Date("2100-01-01")),
           sec.aids.date=as.Date(sec.aids.date, "1970-01-01"),
           death.date=ifelse(year(death.date) >= 1996,
                             as.Date("2100-01-01"),
                             death.date),
           death.date=as.Date(death.date, "1970-01-01")) %>%
    rowwise() %>%
    mutate(censor.date=min(death.date, as.Date("1995-12-31")),
           censor.date=as.Date(censor.date, "1970-01-01")) %>%
    ungroup() %>%
    tbl_df()

@@

tbl.pre <- data %>%
    transmute(site=CENTER,
              race=race,
              age.ge.40=ifelse(age.at.aids >= 40,
                               "age >= 40", "age < 40"),
              aids.year=cut(year(aids.date), c(1984, 1988, 1992, 1995), 
                                 include.lowest=TRUE),
              cd4.base=cut(cd4, c(0, 50, 200, Inf), include.lowest=TRUE),
              initial.aids=frst.aids.diag,
              second.aids=sec.aids.diag,
              depressed=ifelse(depression, "depressed", "not depressed"),
              drug.user=ifelse(drug.use, "Drug User", "Not a Drug User"),
              site=CENTER,
              died=year(death.date) < 2100,
              survival.time=difftime(death.date, aids.date, 
                                     units="weeks") / 52.25 * 12,
              survival.time=as.numeric(survival.time),
              survival.time=ifelse(died, survival.time, NA),
              survival.time.2=difftime(death.date, sec.aids.date,
                                       units="weeks") / 52.25 * 12,
              survival.time.2=as.numeric(survival.time.2),
              survival.time.2=ifelse(died & had.sec, survival.time.2, NA))

>>>>

# read in horiz data with lots of covariates
covariates <- read_csv("data/macswihs_horiz2013_copy-1.csv") %>%
           # date of death
    mutate(death.date=as.Date("1960-01-01") + FRSTDTHD,
           # date of aids diagnosis
           aids.date=as.Date("1960-01-01") + FRSTAIDD,
           # aids interval
#               aids.year=cut(year(aids.date), c(1984, 1988, 1992, 1995), 
#                                  include.lowest=TRUE, right=FALSE),
           aids.interval=cut(year(aids.date), c(1984, 1990, 1993, 1995),
                             include.lowest=TRUE, right=FALSE),
           # date of secondary aids diagnosis
           sec.aids.date=as.Date("1960-01-01") + SECAIDD,
           # sec aids interval
           seca.interval=cut(year(sec.aids.date), c(1984, 1990, 1993, 1995),
                             include.lowest=TRUE, right=FALSE),
           # what was the first aids diagnosis
           frst.aids.diag=plyr::revalue(as.character(AIDSDIAG), map, 
                                        warn_missing=FALSE),
           # what was the first aids diagnosis
           sec.aids.diag=plyr::revalue(as.character(SECAIDSDIAG), map, 
                                       warn_missing=FALSE),
           # did the person have a sec aids diag
           had.sec=year(sec.aids.date) < 1996,
           dob=as.Date("1960-01-01") + DOB,
           age.at.aids=difftime(aids.date, dob, unit="weeks") / 52.25,
           age.at.aids=as.numeric(age.at.aids)) %>%
    select(STUDYID, 
           aids.date,
           aids.interval,
           sec.aids.date,
           seca.interval,
           had.sec,
           frst.aids.diag, 
           sec.aids.diag,
           death.date,
           race=RACE,
           age.at.aids) %>%
    # only keep pts who developed aids before 1996
    filter(year(aids.date) < 1996) %>%
    mutate(aids.year=year(aids.date),
           aids.month=month(aids.date),
           race=ifelse(race == 1, "white", "non-white")) %>%
    rowwise() %>%
    mutate(sec.aids.date=min(sec.aids.date, 
                             as.Date("1995-12-31"))) %>%
    ungroup() %>%
    mutate(sec.aids.date=ifelse(had.sec, sec.aids.date, 
                                as.Date("2100-01-01")),
           sec.aids.date=as.Date(sec.aids.date, "1970-01-01"),
           death.date=ifelse(year(death.date) >= 1996,
                             as.Date("2100-01-01"),
                             death.date),
           death.date=as.Date(death.date, "1970-01-01")) %>%
    rowwise() %>%
    mutate(censor.date=min(death.date, as.Date("1995-12-31")),
           censor.date=as.Date(censor.date, "1970-01-01")) %>%
    ungroup() %>%
    tbl_df()

@@

tbl.pre <- data %>%
    transmute(site=CENTER,
              race=race,
              age.ge.40=ifelse(age.at.aids >= 40,
                               "age >= 40", "age < 40"),
              aids.year=aids.interval,
#               aids.year=cut(year(aids.date), c(1984, 1988, 1992, 1995), 
#                                  include.lowest=TRUE, right=FALSE),
              cd4.base=cut(cd4, c(0, 50, 200, Inf), include.lowest=TRUE,
                           right=FALSE),
              initial.aids=frst.aids.diag,
              second.aids=sec.aids.diag,
              depressed=ifelse(depression, "depressed", "not depressed"),
              drug.user=ifelse(drug.use, "Drug User", "Not a Drug User"),
              site=CENTER,
              died=year(death.date) < 2100,
              survival.time=difftime(death.date, aids.date, 
                                     units="weeks") / 52.25 * 12,
              survival.time=as.numeric(survival.time),
              survival.time=ifelse(died, survival.time, NA),
              survival.time.2=difftime(death.date, sec.aids.date,
                                       units="weeks") / 52.25 * 12,
              survival.time.2=as.numeric(survival.time.2),
              survival.time.2=ifelse(died & had.sec, survival.time.2, NA))
